---
title: "glpD-G3P-glpR, stochastic simulation"
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
---


-----------------   nomenclature

inactive promoter for glpD   PD_R

  active promoter for glpD   PD

inactive promoter for glpK   PK_R

  active promoter for glpK   PK


============ promoter 

R + PD → PD_R                  rp1= kon_p_R * R * PD       (promoter deactivation) 

R + PK → PK_R                  rp2= kon_p_R * R * PK       (promoter deactivation) 

PD_R → R + PD                  rp3= koff_p_R * PD_R

PK_R → R + PK                  rp4= koff_p_R * PK_R

============ transcription

PD → PD + mD         rc1= ktrsc * PD               (transcription by active promoter) 

PK → PK + mK         rc2= ktrsc * PK               (transcription by active promoter) 

PR --> PR + mR       rc5= ktrsc * PR                  (transcription of R)

============= translation 

mD → D                rl1= ktrsl * mD               (translation of D)

mK → K                rl2= ktrsl * mK               (translation of K)

mR ->R                rl3= ktrsl * mR               (translation of R)

============== transportation (newly added) 

null -> G3P           rt2 = ktrsp_G3P * K     (I will play with this)

============ feedback 

G3P + R → G3P_R                rf1= kon_G3P_R * G3P * R 

G3P + PD_R → G3P_R + PD        rf2= kon_G3P_R * G3P * PD_R   (promoter activation) 

G3P + PK_R → G3P_R + PK        rf3= kon_G3P_R * G3P * PK_R   (promoter activation) 

D + G3P  → null                rf4= kon_G3P_D * G3P * D      (negative feedback)

#### ignored ( DHAP → G3P  )

============ degradation 

mD → null              rd1= dRNA * mD             (mRNA degradation)

mK → null              rd2= dRNA * mK

mR -> null             rd3= dRNA * mR

D → null               rd4= dPROTEIN * D            (protein degradation)

K → null              rd5= dPROTEIN * K

R -> null             rd6= dPROTEIN * R

G3P → null            rd7= dG3P * G3P         (metabolite degradation, do we need these?)

============= 

bacteria volume = pi*(0.8/2)^2*3.0 = 1.51 micrometer^3 (rod shape)
number of gene copy  = 1, i.e. PD_R + PD = PK_R + PK = 1


```{r }

rm(list = ls())
library(tidyverse)
library(adaptivetau)

#source("../ebola_cytokineNet/ML_RomyData/multiplot.R")

# Multiple plot function
#
# ggplot objects can be passed in ..., or to plotlist (as a list of ggplot objects)
# - cols:   Number of columns in layout
# - layout: A matrix specifying the layout. If present, 'cols' is ignored.
#
# If the layout is something like matrix(c(1,2,3,3), nrow=2, byrow=TRUE),
# then plot 1 will go in the upper left, 2 will go in the upper right, and
# 3 will go all the way across the bottom.
#
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  library(grid)

  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)

  numPlots = length(plots)

  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                    ncol = cols, nrow = ceiling(numPlots/cols))
  }

 if (numPlots==1) {
    print(plots[[1]])

  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))

    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))

      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}

```


```{r, prepare the parameters }

transitions = list(
    # promoter
    c(PD_R = +1, PD = -1, R = -1),
    c(PK_R = +1, PK = -1, R = -1), 
    c(PD_R = -1, PD = +1, R = +1),
    c(PK_R = -1, PK = +1, R = +1), 
    # transcription
    c(mD = +1),
    c(mK = +1), # transcription by active promoter
    c(mR = +1),
    # translation 
    c(D = +1), 
    c(K = +1),
    c(R = +1),
    # transportation
    c(G3P = +1),
    # feedback
    c(G3P = -1, R = -1),
    c(G3P = -1, PD_R = -1, PD = +1),
    c(G3P = -1, PK_R = -1, PK = +1),
    c(G3P = -1, D = -1), # negative feedback
    # degradation 
    c(mD = -1),
    c(mK = -1),
    c(mR = -1),
    c(D = -1),
    c(K = -1),
    c(R = -1),
    c(G3P = -1)
)

rates <- function(x, params, t) {
    # promoter
    rp1 <- params$kon_p_R * x["R"] * x["PD"]
    rp2 <- params$kon_p_R * x["R"] * x["PK"]
    rp3 <- params$koff_p_R* x["PD_R"]
    rp4 <- params$koff_p_R* x["PK_R"]
    # transcription
    rc1 <- params$ktrsc * x["PD"]
    rc2 <- params$ktrsc * x["PK"]
    rc3 <- params$ktrsc * x["PR"]   / params$r_R
    # translation 
    rl1 <- params$ktrsl * x["mD"]
    rl2 <- params$ktrsl * x["mK"]
    rl3 <- params$ktrsl * x["mR"]
    # transportation
    rt1 <- params$ktrsp_G3P * x["K"]
    # feedback
    rf1 <- params$kon_G3P_R * x["G3P"] * x["R"]
    rf2 <- params$kon_G3P_R * x["G3P"] * x["PD_R"]
    rf3 <- params$kon_G3P_R * x["G3P"] * x["PK_R"]
    rf4 <- params$kon_G3P_D * x["G3P"] * x["D"] 
    # degradation
    rd1 <- params$dRNA * x["mD"]
    rd2 <- params$dRNA * x["mK"]
    rd3 <- params$dRNA * x["mR"]
    rd4 <- params$dPROTEIN * x["D"]
    rd5 <- params$dPROTEIN * x["K"]
    rd6 <- params$dPROTEIN * x["R"]
    rd7 <- params$dG3P     * x["G3P"]

    return(c(rp1, rp2, rp3, rp4,
             rc1, rc2, rc3,
             rl1, rl2, rl3,
             rt1, 
             rf1, rf2, rf3, rf4,
             rd1, rd2, rd3, rd4, rd5, rd6, rd7))

}


ini <- c(
    PD_R = 0,
    PK_R = 0,
    PD = 1,
    PK = 1,
    mD = 5,
    mK = 5,
    mR = 10,
    D = 20,
    K = 20,
    R = 20,
    G3P = 20,
    PR = 1
    )

```

```{r parameters}
params <- list(
  ktrsc = 0.5      , #min-1                           
  r_R = 2,                                      
  ktrsl = 0.3       , #min-1                         
  dRNA = 0.069       , #half life 10 min               
  dPROTEIN = 0.023   , #half life 30 min               
  dG3P =  0.005,                                        
  kon_p_R = 0.144,   # roberts 2011, plos comp bio
  koff_p_R = 0.0378, 
  ktrsp_G3P = 0.08,
  kon_G3P_R= 0.005,
  kon_G3P_D= 0.005
)
```


```{r, simulate}

t.simu <- 6600
t.plot <- 600

# set.seed(12345)
t.start <- Sys.time()
rst.tau <- as_tibble(ssa.adaptivetau(ini, transitions, rates, params, tf = t.simu, tl.params = list(epsilon = 0.05)))
t.tau <- Sys.time() - t.start

# set.seed(12345)
# t.start <- Sys.time()
# rst.exact <- as_tibble(ssa.exact(ini, transitions, rates, params, tf = t.simu ))
# t.exact <- Sys.time() - t.start

itp.tau <- simecol::approxTime(rst.tau, seq(t.plot, t.simu, 1), method = 'constant', rule = 2)  # interpolation on defined time grids
# itp.exact <- simecol::approxTime(rst.exact, seq(t.plot, t.simu, 1), method = 'constant', rule = 2)  # interpolation on defined time grids

itp.tau <- as_tibble(itp.tau)
# sm <- bind_cols( summarise(result, mean = mean(mD)),
#                  summarise(result, sd = sd(mD)),
#                  summarise(result, cv = mean(mD) / sd(mD)),
#                  summarise(result, median = median(mD)),
#                  summarise(result, mad = mad(mD)), 
#                  Gly= params$Gly,
#                  elpT= t.start - t.simu)


```

```{r, include=TRUE}
p1 <- ggplot(itp.tau, mapping = aes(x= time, y= mD)) +
    geom_point()
p2 <- ggplot(itp.tau, mapping = aes(x= mD)) +
    geom_histogram()

p3 <- ggplot(itp.tau, mapping = aes(x= time, y= G3P)) +
    geom_point()
p4 <- ggplot(itp.tau, mapping = aes(x= time, y= D)) +
    geom_point()

p5 <- ggplot(itp.tau, mapping = aes(x= time, y= R)) +
    geom_point()
p6 <- ggplot(itp.tau, mapping = aes(x= time, y= K)) +
    geom_point()

mylayout <- matrix(seq(1,6), ncol = 3, byrow = FALSE)

multiplot(p1, p2, p3, p4, p5, p6, layout = mylayout )

```

