---
  title: "glpD-G3P-glpR, stochastic simulation"
output:
  html_document:
  df_print: paged
pdf_document: default
word_document: default
---
  
  
  -----------------   nomenclature

inactive promoter for glpD   PD_R

active promoter for glpD   PD

inactive promoter for glpK   PK_R

active promoter for glpK   PK


============ promoter 

R + PD → PD_R        rp1= kon_p_R * R * PD       (promoter deactivation) 

R + PK → PK_R        rp2= kon_p_R * R * PK       (promoter deactivation) 

PD_R → R + PD        rp3= koff_p_R * PD_R

PK_R → R + PK        rp4= koff_p_R * PK_R

============ transcription

PD → PD + mD         rc1= ktrsc * PD               (transcription by active promoter) 

PK → PK + mK         rc2= ktrsc * PK               (transcription by active promoter) 

PR --> PR + mR       rc5= ktrsc * PR                  (transcription of R)

============= translation 

mD → D                rl1= ktrsl * mD               (translation of D)

mK → K                rl2= ktrsl * mK               (translation of K)

mR ->R                rl3= ktrsl * mR               (translation of R)

============== transportation (newly added) 

null -> G3P           rt2 = ktrsp_G3P * K          (I will play with this)

============ feedback 

G3P + R → null        rf1= kon_G3P_R * G3P * R 

G3P + PD_R → PD       rf2= kon_G3P_R * G3P * PD_R   (promoter activation) 

G3P + PK_R → PK       rf3= kon_G3P_R * G3P * PK_R   (promoter activation) 

D + G3P  → null       rf4= kon_G3P_D * G3P * D      (negative feedback)

#### ignored ( DHAP → G3P  )

============ degradation 

mD → null            rd1= dRNA * mD             (mRNA degradation)

mK → null            rd2= dRNA * mK

mR -> null           rd3= dRNA * mR

D → null             rd4= dPROTEIN * D            (protein degradation)

K → null             rd5= dPROTEIN * K

R -> null            rd6= dPROTEIN * R

G3P → null           rd7= dG3P * G3P        (metabolite degradation, do we need these?)

============= 
  
  bacteria volume = pi*(0.8/2)^2*3.0 = 1.51 micrometer^3 (rod shape)
number of gene copy  = 1, i.e. PD_R + PD = PK_R + PK = 1


```{r, include=FALSE}

rm(list = ls())
library(tidyverse)
library(adaptivetau)
library(doFuture)
library(DescTools)
library(latex2exp)
registerDoFuture()
options(doFuture.foreach.export = ".export-and-automatic-with-warning")
# plan(multisession(workers = availableCores()/2 - 2) )
# plan(multicore(workers = availableCores()/2 - 2) )

# with multicore, the parallel runs have the same random number seed. 
# with multisession, the seeds are different (associated with more overheads), but not reproducible  

library(doRNG)
# doRNG leads to reproducible, random seed on each slave. 

#source("../ebola_cytokineNet/ML_RomyData/multiplot.R")
# Multiple plot function
#
# ggplot objects can be passed in ..., or to plotlist (as a list of ggplot objects)
# - cols:   Number of columns in layout
# - layout: A matrix specifying the layout. If present, 'cols' is ignored.
#
# If the layout is something like matrix(c(1,2,3,3), nrow=2, byrow=TRUE),
# then plot 1 will go in the upper left, 2 will go in the upper right, and
# 3 will go all the way across the bottom.
#
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  library(grid)
  
  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)
  
  numPlots = length(plots)
  
  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                     ncol = cols, nrow = ceiling(numPlots/cols))
  }
  
  if (numPlots==1) {
    print(plots[[1]])
    
  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))
    
    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))
      
      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}

```


```{r, prepare the parameters}

transitions = list(
  # promoter
  c(PD_R = +1, PD = -1, R = -1),
  c(PD_R = -1, PD = +1, R = +1),
  c(PK_R = +1, PK = -1, R = -1), 
  c(PK_R = -1, PK = +1, R = +1), 
  c(R = -1, PA_R = -1, PR = +1),
  c(R = +1, PA_R = +1, PR = -1),
  # transcription
  c(mD = +1),
  c(mK = +1), # transcription by active promoter
  c(mR = +1),
  # translation 
  c(D = +1), 
  c(K = +1),
  c(R = +1),
  # transportation
  c(G3P = +1),
  # feedback
  c(G3P = -1, R = -1, G3PR = 1),
  c(G3PR = -1, G3P = 1, R = 1),
  c(G3P = -1), # negative feedback by D
  # degradation 
  c(mD = -1),
  c(mK = -1),
  c(mR = -1),
  c(D = -1),
  c(K = -1),
  c(R = -1),
  c(G3P = -1)
)

rates <- function(x, params, t) {
  # promoter
  r1 <- params$p1 * x["R"] * x["PD"]
  r2 <- params$p2* x["PD_R"]
  r3 <- params$p3 * x["R"] * x["PK"]
  r4 <- params$p4* x["PK_R"]
  r5 <- params$p5 * x["R"] * x["PA_R"]
  r6 <- params$p6 * x["PR"]
  # transcription
  r7 <- params$p7 * x["PD"]
  r8 <- params$p8 * x["PK"]
  r9 <- params$p9 * x["PR"]
  # translation 
  r10 <- params$p10 * x["mD"]
  r11 <- params$p11 * x["mK"]
  r12 <- params$p12 * x["mR"]
  # transportation
  r13 <- params$p13 * x["K"]
  # feedback
  r14 <- params$p14 * x["G3P"] * x["R"]
  r15 <- params$p15 * x["G3PR"]
  r16 <- params$p16 * x["G3P"] * x["D"]
  # degradation
  r17 <- params$p17 * x["mD"]
  r18 <- params$p18 * x["mK"]
  r19 <- params$p19 * x["mR"]
  r20 <- params$p20 * x["D"]
  r21 <- params$p21 * x["K"]
  r22 <- params$p22 * x["R"]
  r23 <- params$p23 * x["G3P"]

  return(c(r1, r2, r3, r4, r5, r6, r7, r8, r9, r10,
           r11, r12, r13, r14, r15, r16, r17, r18,
           r19, r20, r21, r22, r23))
  
  
}

```

```{r parameters}
# params <- list(
#   p1=0.5, p2=0.05, p3=0.5, p4=0.05,
#   p5=0.2, p6=0.2, p7=0.5, p8=1, p9=0.5, p10=0.5,
#   p11=0.2, p12=0.005, p13=0, p14=0, p15=0.001, p16=0.069, p17=0.069,
#   p18=0.069, p19=0.023, p20=0.023, p21=0.023, p22=0.035, p23=0.144, p24=0.5)
# params <- list(
#   p1=0.5, p2=0.05, p3=0.5, p4=0.05,
#   p5=0.2, p6=0.2, p7=0.2, p8=0.2, p9=0.2, p10=0.1,
#   p11=0.2, p12=0.005, p13=0, p14=0, p15=0.5, p16=5*0.069, p17=5*0.069,
#   p18=5*0.069, p19=0.023, p20=0.023, p21=0.023, p22=0.023, p23=0.144, p24=0.5)

# params <- list(
#   p1=0.5, p2=0.5, p3=0.5, p4=0.5,
#   p5=0.5, p6=0.5, p7=0.5, p8=0.2, p9=0.2, p10=0.2,
#   p11=0.05, p12=0.1, p13=0, p14=0, p15=0.05, p16=0.069, p17=0.069,
#   p18=0.069, p19=0.023, p20=0.023, p21=0.023, p22=0.023, p23=0.05, p24=0.5, p25=0.001)
params <- list(
  p1=0.5, p2=0.5, p3=0.5, p4=0.5,
  p5=0.5, p6=0.5, p7=0.5, p8=0.5, p9=0.2, p10=0.2,
  p11=0.2, p12=0.2, p13=0.05, p14=0.1, p15=0.001, p16=0.05, p17=0.07,
  p18=0.07, p19=0.07, p20=0.033, p21=0.033, p22=0.033, p23=0.033)


ini_glucose <- c(
  PD_R = 1,
  PK_R = 1,
  PD = 0,
  PK = 0,
  mD = 1,
  mK = 1,
  mR = 5,
  D = 1,
  K = 1,
  R = 20,
  G3P = 1,
  PR = 1,
  PA_R=0,
  G3PR =1
)

ini_glycerol <- c(
  PD_R = 0,
  PK_R = 0,
  PD = 1,
  PK = 1,
  mD = 10,
  mK = 10,
  mR = 2,
  D = 20,
  K = 20,
  R = 5,
  G3P = 10,
  PR = 0,
  PA_R=1,
  G3PR=1
)

ini <- ini_glycerol

```


```{r, simulate}

rst_sum <- tribble(~mean_PD,~mean_PK,~mean_PR,~mean_mD,~mean_mK,~mean_mR,~mean_D, ~mean_K, ~mean_R, ~mean_G3P, ~p13)


t.simu <- 500
t.plot <- 0000

n_rep <- 500 #300
n_scan <- 9 # 6
n_max_para <- 25

plan(multisession(workers = min(n_scan, n_max_para )))

p13_array = seq(from= 0.01, to= 0.3, length.out = n_scan)    
p13_array = c(0.01,0.03,0.05,0.08,0.1,0.15,0.2, 0.3, 0.5)   

y <- vector(mode = "list", length = 1)
y <- foreach(e = 1: n_scan,
             .options.RNG = 12345,
             .export= c('params', 'transitions', 'rates', 'ini',
                        't.plot','t.simu', 'n_rep', 'p11_array',
                        'rst_sum', '%>%'),
             .packages = c('adaptivetau',
                           'tidyverse',
                           'simecol')
) %dorng% {
  print(e)
  p13 <- p13_array[e]
  params$p13 <- p13
  
  for(e in 1:n_rep) {
    
    t.start <- Sys.time()
    rst.tau <- as_tibble(ssa.adaptivetau(ini, transitions, rates, params, tf = t.simu, tl.params = list(epsilon = 0.05)))
    t.tau <- Sys.time() - t.start
    
    # set.seed(12345)
    # t.start <- Sys.time()
    # rst.exact <- as_tibble(ssa.exact(ini, transitions, rates, params, tf = t.simu ))
    # t.exact <- Sys.time() - t.start
    
    itp.tau <- simecol::approxTime(rst.tau, seq(t.plot, t.simu, 1), method = 'constant', rule = 2)  # interpolation on defined time grids
    # itp.exact <- simecol::approxTime(rst.exact, seq(t.plot, t.simu, 1), method = 'constant', rule = 2)  # interpolation on defined time grids
    
    itp.tau <- as_tibble(itp.tau)
    rst_sum <- bind_rows(rst_sum, 
                         tribble(~mean_PD,~mean_PK,~mean_PR,~mean_mD,~mean_mK,~mean_mR,~mean_D, ~mean_K, ~mean_R, ~mean_G3P, ~p13,
                                 mean(itp.tau$PD),mean(itp.tau$PK),mean(itp.tau$PR),mean(itp.tau$mD),   mean(itp.tau$mK), mean(itp.tau$mR),
                                 mean(itp.tau$D), mean(itp.tau$K), mean(itp.tau$R),mean(itp.tau$G3P), p13))
  }  
  rst_sum
  
}

plan(sequential) # This is absolutely necessary in order to release the memory used by the slaves. Otherwise there is no free memory after some time. 
for (e in 1:n_scan) {
  rst_sum <- bind_rows(rst_sum, y[[e]])
}


p4 <- ggplot(rst_sum %>% mutate(p11 = as.factor(p13)), aes(x=p13, y= mean_mD)) + 
  geom_violin(kernel = 'biweight')

p6 <- ggplot(rst_sum %>% mutate(p11 = as.factor(p13)), aes(x=p13, y= mean_D)) + 
  geom_violin(kernel = 'biweight')

p8 <- ggplot(rst_sum %>% mutate(p11 = as.factor(p13)), aes(x=p13, y= mean_mR)) + 
  geom_violin(kernel = 'biweight')

p10 <- ggplot(rst_sum %>% mutate(p11 = as.factor(p13)), aes(x=p13, y= mean_R)) + 
  geom_violin(kernel = 'biweight')


mylayout <- matrix(seq(1,4), ncol = 1, byrow = TRUE)

# multiplot(p4, p55, p6, p77, p8,p9,p10,p11, layout = mylayout )

#ggplot(rst_sum, aes(x=mean_mD, y=mean_D))+geom_line()+facet_wrap(.~p11)

colnames(rst_sum)<-c('PD', 'PK', 'PR',  "mD", 'mK', "mR", "D",
                     "K", "R", "G3P", "p13")
melt_data <- melt(rst_sum, id = 'p13')

f_mean<-function(x){
  # mad(x,constant = 1) #Compute the median absolute deviation
  MeanAD(x, center=mean) # Calculates the mean absolute deviation from a center point, typically the sample mean or the median
}
f_median<-function(x){
  mad(x,constant = 1) #Compute the median absolute deviation
  # MeanAD(x, center=mean) # Calculates the mean absolute deviation from a center point, typically the sample mean or the median
}
f_cv<-function(x){
  sd(x)/mean(x)
}
stat_sum_single <- function(fun, color, geom="point", ...) {
  stat_summary(fun.y=fun, colour=color, geom=geom, position = position_dodge(width = 0.9), size = 3, ...)
}
  
p_p<-melt_data[which(melt_data$variable %in% c('R', 'D')),] %>% ggplot(aes(x=as.factor(p13),y=value, fill=variable))+geom_violin(trim = TRUE, position="dodge", alpha=0.5,scale = "width")+xlab(TeX('$k_{ing3p}$')) +
  ylab("Value") +stat_summary(fun.data=mean_sdl, fun.args = list(mult = 1),
               geom="pointrange", color="black",
               shape = 18, size = 0.75,position = position_dodge(width = 0.9))+stat_sum_single(f_mean,'red')+theme_bw()+theme(strip.background=element_rect(fill='white')) +theme(
      panel.background = element_rect(fill = "white",colour = "black",linewidth = 1),panel.grid.major = element_blank(),panel.grid.minor = element_blank(),legend.key.size = unit(1, 'cm'),legend.title = element_blank(),
      legend.text = element_text(colour="black", size=15),legend.position = c(0.05, 0.8),
      axis.text.x = element_text(face="bold", size=12,angle=45, vjust = 0.7),
      axis.text.y = element_text(face="bold", size=12),
      axis.title.x = element_text(face="bold", size=16),
      axis.title.y = element_text(face="bold", size=16))

p_m<-melt_data[which(melt_data$variable %in% c('mR', 'mD')),] %>% ggplot(aes(x=as.factor(p13),y=value, fill=variable))+geom_violin(trim = TRUE, position="dodge", alpha=0.5,scale = "width")+xlab(TeX('$k_{ing3p}$')) +
  ylab("Value") +stat_summary(fun.data=mean_sdl, fun.args = list(mult = 1),
               geom="pointrange", color="black",
               shape = 18, size = 0.75,position = position_dodge(width = 0.9))+stat_sum_single(f_mean,'red')+theme_bw()+theme(strip.background=element_rect(fill='white')) +theme(
      panel.background = element_rect(fill = "white",colour = "black",linewidth = 1),panel.grid.major = element_blank(),panel.grid.minor = element_blank(),legend.key.size = unit(1, 'cm'),legend.title = element_blank(),
      legend.text = element_text(colour="black", size=15),
      legend.position = c(0.05, 0.8),
      axis.text.x = element_text(face="bold", size=12,angle=45, vjust = 0.7),
      axis.text.y = element_text(face="bold", size=12),
      axis.title.x = element_text(face="bold", size=16),
      axis.title.y = element_text(face="bold", size=16))

mylayout <- matrix(seq(1,2), ncol = 1, byrow = TRUE)

pdf('gillespie.pdf', width = 18, height = 12)
multiplot(p_m,p_p, layout = mylayout )
dev.off()



kk<-melt_data[which(melt_data$variable %in% c( 'mean_D') & melt_data$p11 %in% c( '0.01')),]

```

```{r, plot cases}

params$p11 <- 0.05

t.start <- Sys.time()
rst.tau <- as_tibble(ssa.adaptivetau(ini, transitions, rates, params, tf = t.simu, tl.params = list(epsilon = 0.05)))
t.tau <- Sys.time() - t.start

# set.seed(12345)
# t.start <- Sys.time()
# rst.exact <- as_tibble(ssa.exact(ini, transitions, rates, params, tf = t.simu ))
# t.exact <- Sys.time() - t.start

itp.tau <- simecol::approxTime(rst.tau, seq(t.plot, t.simu, 1), method = 'constant', rule = 2)  # interpolation on defined time grids
# itp.exact <- simecol::approxTime(rst.exact, seq(t.plot, t.simu, 1), method = 'constant', rule = 2)  # interpolation on defined time grids

itp.tau <- as_tibble(itp.tau)

p1 <- ggplot(itp.tau, mapping = aes(x= time, y= mD)) +
  geom_point()
p2 <- ggplot(itp.tau, mapping = aes(x= mD)) +
  geom_histogram()

p3 <- ggplot(itp.tau, mapping = aes(x= time, y= G3P)) +
  geom_point()
p4 <- ggplot(itp.tau, mapping = aes(x= time, y= R)) +
  geom_point()

p5 <- ggplot(itp.tau, mapping = aes(x= time, y= D)) +
  geom_point()
p6 <- ggplot(itp.tau, mapping = aes(x= time, y= K)) +
  geom_point()

mylayout <- matrix(seq(1,4), ncol = 2, byrow = FALSE)

multiplot(p1, p2, p3, p4, layout = mylayout )

```

